Wikipedia Text Mining and Entropy Analysis

### 2019-01-25 cogsci/position-info.Rmd
almost done with second cogsci draft, working on imputation and glm

@@

@@

We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of telephone conversations [CITATION].

We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages. Children are not fully developed speakers, and so we expect to observe a different shape of the distribution. We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position.

@@

## Results & Analysis
The adult and child entropy curves track one another almost identically [Make this contrast with the prediction stronger, introduce . This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence. We attribute the penultimate dip in these unigram entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances. The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position. The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993). The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.

# Written data
We also wanted to apply this analysis to large-scale text data, to compare with the parent-child speech data results from using CHILDES. WHY ARE WE DOING THIS.

@@

To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013). We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia. [Miss MDA or do another kind of missing value imputation, description of method]

@@

## Flora's feedback: 
1. Extend the background section with more information about the literature (what are some of the consequence of the unigram entropy, where has this been applied). 
2. Talk about the speech and text distinction in the literature and how that affects your work. 
3. Talk about how your work looks at multiple languages (has this been done before). 
4. Talk more about child versus adult distinction. 
6. Talk more about entropy in general, and then why this measure applies to this work (Genzel and Charniak?). 
7. Have a method and results section (i.e. clearer division) with both the CHILDES and the Wikipedia data.
## Self feedback:
1. Motivating sections and why you chose the corpora you did
2. Clean up and rerun Wikipedia code
3. Write WALS features section. Why do we care about the typological features? 
4. Write clustering section and section about what features matter explored using statistical modeling. 
5. Spinal column: Why does information structure through entropy matter? How has it impacted the literature (Genzel & Charniak, Hale, Levy, Piantadosi, etc.)? Why does UID matter? What does the stuff we do here show? Why do we choose the particular corpora, entropy methods and analyses we do? How does this potentially connect with eye-tracking and sentence processing in general?

>>>>

UID has been applied broadly over the past decade, to determining whether linguistic alignment takes place (Jaeger & Snider, 2013), Zipfian word length distributions (Piantadosi et al., 2011), communication efficiency (Mahowald et al., 2013), dialogue and turn-taking (Xu & Reitter, 2018) and the significance of ambiguity in language (Piantadosi et al., 2012), among other research.

@@

The UID hypothesis predicts that information transmission rate will tend towards uniformity in both speech and text. We begin by examining speech to determine if we obtain the entropy curve shape we expect for spoken communication. We begin with the same source as Levy & Jaeger (2007) and Jaeger (2010), the Switchboard corpus of adult telephone conversations [CITATION]. We also examine child-adult conversations in the CHILDES TalkBank (Brown 1973; MacWhinney 2000) corpora database of spoken adult-child conversations in multiple languages. Children are not fully developed speakers, so we want to compare the entropy curve we obtain by computing over the utterances in the CHILDES corpora to the utterances in the adult-adult Switchboard corpus.

@@

Switchboard

We first used the Brown and Providence English corpora from CHILDES. The Brown corpus contains individual recordings of conversations between three children between 1.5 and 6-years-old and their families in their homes. The Providence corpus recorded interactions between children between 1 and 3 years old and their parents in the home. We divided each corpus by speaker into child and non-child categories. We further divided the corpora by utterance length, so that all sentences of length $k$ (e.g. $6$) were grouped together. Finally, within each utterance length, we computed the unigram entropy measure for each position.

@@

## Results & Analysis
The adult and child entropy curves track one another almost identically. This is surprising because UID would predict that the young children in the corpora who are not fully developed speakers would have more noise in their distribution of information rates. This not only indicates a robustness in the unigram entropy curve across speakers, but also across ages and addressees. We also observed a robustness across corpora for the same languages, and a robustness across utterance lengths within the same corpus's entropy curve. This shows that the concept of an "entropy curve" for a specific language is well-founded when considering speech data.

We found a distinct three-step distribution for English and Spanish CHILDES corpora, with a slight dip in the penultimate position of each sentence. The final position of utterances in child-directed speech is known to be important, dating back to Aslin (1993). The Mandarin corpus entropy curve, by comparison, has a noticeably lower positional entropy values in utterance-final positions than in utterance-penultimate positions.

We attribute the penultimate dip in the English and Spanish entropy curves to the fact that most of the utterances in the CHILDES English and Spanish corpora we examined had a determiner such as "the" or "a" in the second-to-last position of utterances. The beginnings of utterances in the English and Spanish CHILDES data were usually pronouns or grammatical subjects, while the final words were grammatical objects and had a great deal of variation in the exact word that appeared in the utterance-final position.

These entropy curves are not what would be expected from UID. The robustness of the three-step distribution for English and its replication in Spanish do not resemble the affine function we would expect from UID. The Mandarin entropy curve, which does not at all resemble either the English/Spanish distribution or the predictions of UID, suggests that the entropy curve can vary from language to language. UID predicts that each language should have a similar distribution.

# Written data

The UID hypothesis also applies to written communication: we expect people to communicate information at a uniform rate through writing as well. We use Wikipedia as a source for written data, which provides two advantages. One, the quantity of data in Wikipedia is large for each language and two, there are hundreds of languages with Wikipedias. This allows us to perform the entropy analysis on a much greater scale and to directly compare the results of the entropy analysis on each language to one another, and ultimately to predict what typological features of a language help determine its entropy curve, if any. We will describe our method of harvesting and distilling text data from each Wikipedia as well as how we compare the entropy curves from each language to one another.

@@

To determine which phonological, morphological and syntactic features affected the embedding of a language in the Wikipedia dataset, we looked at the list of $144$ linguistic features in World Atlas of Language Structures (Dryer & Haspelmath, 2013). We limited the languages in the WALS database to only those in our Wikipedia dataset and performed missing-value imputation to obtain the features not coded in WALS for the languages from Wikipedia.

@@

### 2019-01-25 wiki_entropy/wiki_clustering.Rmd
almost done with second cogsci draft, working on imputation and glm

```{r clustering}
NUM_CLUSTERS <- 10

@@

```{r MCA_imputation}
wiki_gs <- gs_title("Wikipedia_absolute")
wkpd <- gs_read(wiki_gs) 
LANGUAGES <- wkpd$Language

for (i in 1:4) {
  wf.sub[, i] <- factor(wf.sub[, i])
}

wf.sub <- wf.sub[rowSums(is.na(wf.sub)) != 4, ]
rownames(wf.sub) <- wf.sub$language
wf.sub$language <- NULL
rv <- MIMCA(wf.sub, ncp = 4)

imputed_wf <- as.data.frame(rv[[1]][[99]])

wals_clust_full <- kmodes(imputed_wf, NUM_CLUSTERS, iter.max=10, weighted = FALSE)
```

@@

>>>>

```{r hierarchical clustering}
NUM_CLUSTERS <- 10

@@

@@

```{r imputation for glm}
imputed_result <- wf_df %>%
  select(language, feature, value) %>%
  ungroup() %>% 
  spread(feature, value) %>% 
  column_to_rownames("language") %>%
  sapply(as.factor) %>%
  as_data_frame() %>%
  MIMCA(ncp = 7, maxiter = 999)

imputed_wf <- as.data.frame(imputed_result[[1]][[99]])
```
